\setcounter{subsection}{3-1}
\subsection{Relations}

% Some useful commands for these exercises
\newcommand\ppt[1]{(x_{#1}, y_{#1})}
\newcommand\peq[1]{y_{#1} - x_{#1}^2}

\exercise{1}{
  Define two points $\ppt{0}$ and $\ppt{1}$ of the plane to be equivalent if $\peq{0} = \peq{1}$.
  Check that this is an equivalence relation and describe the equivalence classes.
}
\sol{
  First we show that this relation, which we shall denote with $\sim$, is an equivalence relation.
  \qproof{
    In what follows, suppose that $\ppt{0}$, $\ppt{1}$, and $\ppt{2}$ are all points in the plane.
    
    (Reflexivity) Of course we have $\peq{0} = \peq{0}$, and hence $\ppt{0} \sim \ppt{0}$.

    (Symmetry) Suppose that $\ppt{0} \sim \ppt{1}$.
    Then we have $\peq{0} = \peq{1}$ so that of course $\peq{1} = \peq{1}$ since numerical equality is symmetric, and so $\ppt{1} \sim \ppt{0}$ as well.

    (Transitivity) Suppose that $\ppt{0} \sim \ppt{1}$ and $\ppt{1} \sim \ppt{2}$.
    Then $\peq{0} = \peq{1}$ and $\peq{1} = \peq{2}$ so that of course $\peq{0} = \peq{2}$ since numerical equality is transitive.
    Therefore $\ppt{0} \sim \ppt{2}$, which shows transitivity.

    This suffices to show that $\sim$ is an equivalence relation as we set out to show.
  }

  Each equivalence class formed by this relation is the parabola $y = x^2$ shifted up or down on the $y$-axis.
  This is easy to see since two points $\ppt{0}$ and $\ppt{1}$ are in the same class if $\peq{0}$ and $\peq{1}$ have the same value, say $c$.
  Then $\peq{0} = c$ so that $y_0 = x_0^2 + c$, which is clearly such a parabola, and similarly $y_1 = x_1^2 + c$.
}

\exercise{2}{
  Let $C$ be a relation on a set $A$.
  If $A_0 \ss A$, define the \boldit{restriction} of $C$ to $A_0$ to be the relation $C \cap (A_0 \times A_0)$.
  We also note that clearly $C_0 \ss C$ as well.
  Show that the restriction of an equivalence relation is an equivalence relation.
}
\sol{
  \qproof{
    Define $C$, $A$, and $A_0$ as above and suppose that $C$ is an equivalence relation.
    Let $C_0 = C \cap (A_0 \times A_0)$ be the restriction of $C$ to $A_0$, noting that this is in fact a relation on $A_0$ since clearly $C_0 \ss A_0 \times A_0$.
    Now we show that $C_0$ satisfies the three properties of an equivalence relation.

    (Reflexivity) Consider any $a \in A_0$ so that of course $(a,a) \in A_0 \times A_0$.
    Since $A_0 \ss A$ we also have that $a \in A$.
    Hence $aCa$ since $C$ is an equivalence relation on $A$ and is therefore reflexive.
    Thus $(a,a) \in C \cap (A_0 \times A_0) = C_0$, which shows that $a C_0 a$ so that $C_0$ is reflexive since $a$ was arbitrary.

    (Symmetry) Suppose that $a,b \in A_0$ and that $aC_0b$.
    Then of course $(b,a) \in A_0 \times A_0$ and $bCa$ since $C_0 \ss C$.
    From this it follows that $(b,a) \in C \cap (A_0 \times A_0) = C_0$ so that $b C_0 a$.
    This of course shows that $C_0$ is symmetric.

    (Transitivity) Now consider $a,b,c \in A_0$ and suppose that both $a C_0 b$ and $b C_0 c$.
    Then we have $aCb$ and $bCc$ since $C_0 \ss C$.
    Since $C$ is an equivalence relation and therefore transitive, it follows that $aCc$, and since also clearly $(a,c) \in A_0 \times A_0$, we have $(a,c) \in C \cap (A_0 \times A_0) = C_0$ so that $a C_0 c$.
    This shows that $C_0$ is transitive.
  }
}

\exercise{3}{
  Here is a ``proof'' that every relation $C$ that is both symmetric and transitive is also reflexive:
  ``Since $C$ is symmetric, $aCb$ implies $bCa$.
  Since $C$ is transitive, $aCb$ and $bCa$ together imply $aCa$, as desired.''
  Find the flaw in this argument.
}
\sol{
  Suppose that $C$ is a relation on the set $A$.
  This argument is perfectly valid for any $a,b \in A$ such that $aCb$, which is to say that we can conclude that $aCa$ in this case (and by the same argument $bCb$).
  However, reflexivity requires $aCa$ to hold for \emph{every} $a \in A$.
  So if there is no $b \in A$ such that $aCb$ then the above argument cannot be applied and we cannot conclude that $aCa$.
  In this case the element $a$ is effectively not involved in the relation at all.

  This is perhaps best illustrated with an example: suppose that $A = \braces{1,2,3,4}$ and
  \gath{
    C = \braces{(1,1), (2,2), (3,3), (1,2), (2,1), (2,3), (3,2), (1,3), (3,1)} \,.
  }
  It is easy to verify that $C$ is both symmetric and transitive on $A$ but it is clearly not reflexive since $(4,4) \notin C$.
  One can also observe how $4$ is not involved in the relation at all and, if it were, it would have to be that $(4,4) \in C$ if $C$ were to remain symmetric and transitive.
}

\exercise{4}{
  Let $f: A \to B$ be a surjective function.
  Let us define a relation on $A$ by setting $a_0 \sim a_1$ if
  \gath{
    f(a_0) = f(a_1) \,.
  }
  \eparts{
  \item Show that this is an equivalence relation.
  \item Let $A^*$ be the set of equivalence classes.
    Show that there is a bijective correspondence of $A^*$ with $B$.
  }
}
\sol{
  (a)
  \qproof{
    We show the three properties necessary for $\sim$ to be an equivalence relation:
    
    (Reflexivity) Consider any $a \in A$ so that of course $f(a) = f(a)$ since $f$ is a function.
    Hence $a \sim a$ so that $\sim$ is reflexive since $a$ was arbitrary.

    (Symmetry) Consider $a,b \in A$ and suppose that $a \sim b$.
    Then by definition $f(a) = f(b)$ so that obviously also $f(b) = f(a)$ since equality is symmetric.
    So of course $b \sim a$, which shows that $\sim$ is symmetric.

    (Transitivity) Consider $a,b,c \in A$ and suppose that $a \sim b$ and $b \sim c$.
    Then by definition $f(a) = f(b)$ and $f(b) = f(c)$ so that of course $f(a) = f(b) = f(c)$, and hence $a \sim c$.
    This shows that $\sim$ is transitive.
  }

  (b)
  \qproof{
    Define the function $g: A^* \to B$ as follows.
    For any equivalence class $C \in A^*$, we know that $C$ is nonempty since $A^*$ is a partition of $A$.
    Hence there is an $a \in C$, so set $g(C) = f(a)$, noting that clearly $g(C) = f(a) \in B$ so that $B$ can be the range of $g$.

    To show that $g$ is injective, consider two equivalence classes $C$ and $D$ where $g(C) = g(D)$.
    Then there are elements $c \in C$ and $d \in D$ where $f(c) = g(C) = g(D) = f(d)$.
    This shows that $c \sim d$ so that they must be in the same equivalence class.
    Thus $d \in C$ since $c \in C$, but also $d \in D$ so that $C$ and $D$ are not disjoint.
    Hence it must be that $C = D$ by Lemma~3.1, which shows that $g$ is injective.

    To show that $g$ is surjective, consider any $b \in B$.
    Since $f$ is surjective, there is an $a \in A$ such that $f(a) = b$.
    Since $A^*$ is a partition, $a$ must belong to an equivalence class $C \in A^*$.
    Then there is an element $c \in C$ such that $g(C) = f(c)$ by the definition of $g$.
    Since $a$ and $c$ are both in the same equivalence class $C$, we have that $a \sim c$ so that $g(C) = f(c) = f(a) = b$.
    This shows that $g$ is surjective since $b \in B$ was arbitrary.

    Therefore we have shown that $g$ is both injective and surjective, and so is a bijection by definition, as desired.
  }
}

\exercise{5}{
  Let $S$ and $S'$ be the following subsets of the plane:
  \ali{
    S &= \braces{\ppt{} \where y = x+1 \text{ and } 0 < x < 2} \,, \\
    S' &= \braces{\ppt{} \where y - x \text{ is an integer}} \,.
  }
  \eparts{
  \item Show that $S'$ is an equivalence relation on the real line and $S' \sps S$.
    Describe the equivalence classes of $S'$.
  \item Show that given any collection of equivalence relations on a set $A$, their intersection is an equivalence relation on $A$.
  \item Describe the equivalence relation $T$ on the real line that is the intersection of all equivalence relations on the real line that contain $S$.
    Describe the equivalence classes of $T$.
  }
}
\sol{
  (a)
  \qproof{
    First note that $S' \ss \reals \times \reals$ and so is a relation on $\reals$.
    We show that $S'$ has the three properties required of an equivalence relation.

    (Reflexivity) Consider any $x \in \reals$ so that clearly $x - x=0$ is an integer.
    Hence $(x,x) \in S'$ by definition.
    This shows that $S'$ is reflexive since $x$ was arbitrary.

    (Symmetry) Suppose that $x,y \in \reals$ and $x S' y$.
    Then $n = y - x$ is an integer so that $x - y = -(y-x) = -n$ is also clearly an integer.
    Therefore $y S' x$ as well, which shows that $S'$ is symmetric.

    (Transitivity) Consider $x,y,z \in \reals$ and suppose that both $x S' y$ and $y S' z$.
    Then $n = y - x$ and $m = z - y$ are both integers.
    We then have
    \gath{
      z - x = z - x + y - y = (z - y) + (y - x) = m + n \,,
    }
    which is clearly an integer since $m$ and $n$ are.
    Hence $x S' z$ so that $S'$ is transitive.

    It is easy to show that $S' \sps S$.
    Consider any $\ppt{} \in S$ so that $0 < x < 2$ and $y = x + 1$.
    Then $y - x = (x + 1) - x = 1$, which is of course an integer.
    Hence $\ppt{} \in S'$, and thus $S' \sps S$ since $\ppt{}$ was arbitrary.
  }

  The equivalence class $C$ containing $x \in \reals$ is the countable set $C = \braces{x + n \where n \in \ints}$.
  While perhaps not immediately obvious, it is almost trivial to show:
  \ali{
    y \in C &\bic \exists n \in \ints (y = x + n) \bic \exists n \in \ints (y - x = n) \\
    &\bic x S' y \bic y S' x \\
    &\bic \text{$y$ is in the equivalence class determined by $x$}
  }
  since $S'$ is symmetric.

  (b)
  \qproof{
    Let $A^*$ be a collection of equivalence relations on $A$ so that we must show that $C = \bigcap_{D \in A^*} D$ is also an equivalence relation on $A$.
    First, suppose that any $\ppt{} \in C$ and consider any $D \in A^*$ so that $\ppt{} \in D$.
    Then also $\ppt{} \in A \times A$ since $D$ is a relation on $A$ so that $D \ss A \times A$.
    This shows that $C \ss A \times A$ since $\ppt{}$ was arbitrary, and so $C$ is a relation on $A$.
    Now we show the three required properties of an equivalence relation:

    (Reflexivity) Consider any $x \in A$ so that $(x,x) \in D$ for every $D \in A^*$ since each $D$ is an equivalence relation and so is reflexive.
    It then follows that $(x,x) \bigcap_{D \in A^*} D = C$, which shows that $C$ is reflexive.

    (Symmetric) Suppose that $\ppt{} \in C$ and consider any $D \in A^*$ so that also $\ppt{} \in D$.
    Then also $(y,x) \in D$ since $D$ is an equivalence relation and so is symmetric.
    Since $D$ was arbitrary, this shows that $(y,x) \in \bigcap_{D \in A^*} D = C$ so that $C$ is symmetric.

    (Transitivity) Suppose that $\ppt{} \in C$ and $(y,z) \in C$.
    For any $D \in A^*$ we then have that both $\ppt{} \in D$ and $(y,z) \in D$.
    It then follows that $(x,z) \in D$ since $D$ is an equivalence relation and so is transitive.
    Since $D$ was arbitrary, we have that $(x,z) \in \bigcap_{D \in A^*} D = C$ so that $C$ is transitive as desired.
  }

  (c)
  First we note that $S$ itself is \emph{not} an equivalence relation on $\reals$ since it is not reflexive.
  In fact $(x,x) \notin S$ for any $x \in \reals$ since it is never true that $x = x+1$.
  Now define the following subsets of the plane:
  \ali{
    S_1 &= \braces{\ppt{} \where y = x} &
    S_4 &= \braces{\ppt{} \where y = x + 2 \text{ and } 0 < x < 1} \\
    S_2 &= S = \braces{\ppt{} \where y = x + 1 \text{ and } 0 < x < 2} &
    S_5 &= \braces{\ppt{} \where y = x - 2 \text{ and } 2 < x < 3} \,. \\
    S_3 &= \braces{\ppt{} \where y = x - 1 \text{ and } 1 < x < 3}
  }
  We then claim that the intersection we seek is $T = \bigcup_{i \in \intsfin{5}} S_i$.
  An illustration of this set in the plane is shown below:
  \begin{center}
    \begin{tikzpicture}[scale=0.6]
      % Open line
      \newcommand\oline[2]{
        \draw[thick,-] (#1) -- (#2);
        \filldraw[fill=white] (#1) circle (0.15);
        \filldraw[fill=white] (#2) circle (0.15);
      }
      % Coordinate grid
      \draw[step=1cm,black!25!white,thin] (-0.9,-0.9) grid (3.9,3.9);

      % Axes
      \draw[thick,<->] (-1,0) -- (4,0);
      \draw[thick,<->] (0,-1) -- (0,4);

      % Subsets
      \draw[thick,<->] (-0.9,-0.9) -- (3.9,3.9);
      \node [below] at (3.5,3.5) {$S_1$};
      \oline{0,1}{2,3}
      \node [above] at (2,3) {$S_2$};
      \oline{1,0}{3,2}
      \node [right] at (3,2) {$S_3$};
      \oline{0,2}{1,3}
      \node [above] at (1,3) {$S_4$};
      \oline{2,0}{3,1}
      \node [right] at (3,1) {$S_5$};
    \end{tikzpicture}
  \end{center}
  \qproof{
    Let $S^*$ denote the collection of all equivalence relations on $\reals$ that contain $S$ so that we must show that $T = \bigcap_{R \in S^*} R$.

    $(\ss)$ Consider any $\ppt{} \in T$ and any $R \in S^*$ so that $R$ is an equivalence relation on $\reals$ that contains $S$.
    We then have the following cases:

    Case: $\ppt{} \in S_1$.
    Then of course $y = x$ so that $\ppt{} = (x,x) \in R$ since it is an equivalence relation and thus reflexive.

    Case: $\ppt{} \in S_2 = S$.
    Then of course $\ppt{} \in R$ since $R$ contains $S$.

    Case: $\ppt{} \in S_3$.
    Then we have that $y = x - 1$ with $1 < x < 3$, from which it follows that $x = y + 1$ and $0 < y = x-1 < 2$.
    Therefore $(y,x) \in S$ so that $(y,x) \in R$ since $R$ contains $S$.
    We then have that $\ppt{} \in R$ as well since $R$ is an equivalence relation and therefore symmetric.

    Case: $\ppt{} \in S_4$.
    Then $y = x+2$ with $0 < x < 1$.
    Let $z = x+1$ so that $(x,z) \in S = S_4$ since we also have $0 < x < 1 < 2$.
    We then know that $(x,z) \in R$ since $R$ contains $S$.
    We also then have that $y = x+2 = (x+1)+1 = z+1$ with $0 < 1 < z = x+1 < 2$ since $0 < x < 1$.
    Thus $(z,y) \in S = S_4$ so that also $(z,y) \in R$.
    Since $R$ is an equivalence relation and therefore transitive, we have that $\ppt{} \in R$ as well.

    Case: $\ppt{} \in S_5$.
    Then we have $y = x-2$ and $2 < x < 3$.
    It then follows that $x = y+2$ and $0 < y = x-2 < 1$ so that $(y,x) \in S_4$.
    Then of course $(y,x) \in R$ by the previous case so that also $\ppt{} \in R$ since $R$ is an equivalence relation and therefore symmetric.

    Thus in every case $\ppt{} \in R$ so that also $\ppt{} \in \bigcap_{R \in S^*} R$ since $R$ was arbitrary.
    It then follows that $T \ss \bigcap_{R \in S^*} R$ since $\ppt{}$ was arbitrary.

    $(\sps)$ All we really need to show is that $T$ is an equivalence relation on $\reals$ that contains $S$.
    From this it follows that $T \in S^*$ so that of course $T \sps \bigcap_{R \in S^*} R$.
    First we note that clearly $T \ss \reals^2$ so that it is a relation on $\reals$.
    Also clearly it contains $S$ since $S = S_2 \ss T$.
    Now we show that it has the three properties that an equivalence relation require.

    (Reflexivity) For any $x \in \reals$ clearly $(x,x) \in S_1 \ss T$ so that $T$ is reflexive.

    (Symmetry) Suppose that $xTy$ so that $\ppt{} \in T$.
    We then have the following cases:

    Case: $\ppt{} \in S_1$.
    Then of course $y = x$ so $(y,x) = (x,x) = \ppt{} \in S_1 \ss T$.

    Case: $\ppt{} \in S_2 = S$.
    Then $y = x+1$ and $0<x<2$ so that $x = y-1$ and $1 < y=x+1 < 3$.
    Hence $(y,x) \in S_3 \ss T$.

    Case: $\ppt{} \in S_3$.
    Then we have that $y = x - 1$ with $1 < x < 3$, from which it follows that $x = y + 1$ and $0 < y = x-1 < 2$.
    Therefore $(y,x) \in S = S_2 \ss T$.

    Case: $\ppt{} \in S_4$.
    Then $y = x+2$ with $0 < x < 1$ so that $x = y-2$ and $2 < y=x+2 < 3$.
    Hence $(y,x) \in S_5 \ss T$.

    Case: $\ppt{} \in S_5$.
    Then we have $y = x-2$ and $2 < x < 3$ so that $x = y+2$ and $0 < y=x-2 < 1$.
    Hence $(y,x) \in S_4 \ss T$.

    So in all cases $(y,x) \in T$, which shows that $T$ is symmetric.

    (Transitivity) Now suppose that $xTy$ and $yTz$.
    If $x = y$ then of course we have $(x,z) = (y,z) \in T$.
    Similarly if $y = z$ then $(x,z) = \ppt{} \in T$.
    So assume that $x \neq y$ and $y \neq z$ so that it can neither be that $\ppt{} \in S_1$ nor $(y,z) \in S_1$.
    Thus there are four sets (i.e. $S_i$ where $i \in \braces{2,3,4,5}$) that $\ppt{}$ and $(y,z)$ can be in, which results in sixteen different possibilities, though not all are possible:

    Case: $\ppt{} \in S_2$.
    Then $y=x+1$ and $0 < x < 2$ so that $1 < y = x+1 < 3$.
    \begin{indpar}
      Case: $(y,z) \in S_2$.
      Then also $z=y+1$ and $0 < y < 2$ so that $z = y+1 = (x+1)+1 = x+2$ and $y = x+1 < 2$ means that $x < 1$.
      Hence $z = x+2$ and $0 < x < 1$ so that $(x,z) \in S_4 \ss T$.

      Case: $(y,z) \in S_3$.
      Then $z = y-1$ and $1 < y < 3$ so that $z = y-1 = (x+1)-1 = x$, and hence $(x,z) = (x,x) \in S_1 \ss T$.

      Case: $(y,z) \in S_4$.
      This case is not possible because $1 < y < 3$ so that it cannot be that $0 < y < 1$ and hence $(y,z)$ cannot be in $S_4$.

      Case: $(y,z) \in S_5$.
      Then $z = y-2$ and $2 < y < 3$ so that $z = y-2 = (x+1)-2 = x-1$ and $2 < y = x+1 < 3$ means that $1 < x < 2 < 3$.
      Hence $z=x-1$ and $1 < x < 3$ so that $(x,z) \in S_3 \ss T$.
    \end{indpar}

    Case: $\ppt{} \in S_3$.
    Then $y=x-1$ and $1 < x < 3$ so that $0 < y = x-1 < 2$.
    \begin{indpar}
      Case: $(y,z) \in S_2$.
      Then $z = y+1$ and $0 < y < 2$ so that $z = y+1 = (x-1)+1 = x$ and hence $(x,z) = (x,x) \in S_1 \ss T$.

      Case: $(y,z) \in S_3$.
      Then $z = y-1$ and $1 < y < 3$ so that $z = y-1 = (x-1)-1 = x-2$ and $1 < y = x-1$ and hence $2 < x$.
      Therefore $z = x-2$ and $2 < x < 3$ so that $(x,z) \in S_5 \ss T$.

      Case: $(y,z) \in S_4$.
      Then $z=y+2$ and $0 < y < 1$ so that $z = y+2 = (x-1)+2 = x+1$ and $y = x-1 < 1$ and hence $x < 2$.
      Thus $z = x+1$ and $0 < 1 < x < 2$ so that $(x,z) \in S_2 \ss T$.

      Case: $(y,z) \in S_5$.
      This is case is not possible because $y < 2$ so that it cannot be that $2 < y < 3$, and hence $(y,z)$ cannot be in $S_5$.
    \end{indpar}

    Case: $\ppt{} \in S_4$.
    Then $y=x+2$ and $0 < x < 1$ so that $2 < y = x+2 < 3$.
    \begin{indpar}
      Case: $(y,z) \in S_2$.
      This case is also impossible because $2 < y$  so that it cannot be that $0 < y < 2$, and hence $(y,z)$ cannot be in $S_2$.

      Case: $(y,z) \in S_3$.
      Then $z = y-1$ and $1 < y < 3$ so that $z = y-1 = (x+2)-1 = x+1$ and $y = x+2 < 3$ so that $x < 1 < 2$.
      Thus $z = x+1$ and $0 < x < 2$ so that $(x,z) \in S_2 \ss T$.

      Case: $(y,z) \in S_4$.
      This case is not possible because again $2 < y$ so that it cannot be that $0 < y < 1$, and hence $(y,z)$ cannot be in $S_4$.

      Case: $(y,z) \in S_5$.
      Then $z=y-2$ and $0 < y <1$ so that $z=y-2=(x+2)-2=x$ and hence $(x,z) = (x,x) \in S_1 \ss T$.
    \end{indpar}

    Case: $\ppt{} \in S_5$.
    Then $y=x-2$ and $2 < x < 3$ so that $0 < y=x-2 < 1$.
    \begin{indpar}
      Case: $(y,z) \in S_2$.
      Then $z = y+1$ and $0 < y < 2$ so that $z = y+1 = (x-2) + 1 = x-1$ and $0 < y = x-2$ so that $1 < 2 < x$.
      Therefore $z=x-1$ and $1 < x < 3$ so that $(x,z) \in S_3 \ss T$.

      Case: $(y,z) \in S_3$.
      This case is not possible because $y < 1$ so that it cannot be that $1 < y < 3$, and hence $(y,z)$ cannot be in $S_3$.

      Case: $(y,z) \in S_4$.
      Then $z = y+2$ and $0 < y < 1$ so that $z = y+2 = (x-2)+2$ and hence $(z,x) = (x,x) \in S_1 \ss T$.

      Case: $(y,z) \in S_5$.
      This case is also not possible because again $y < 1$ so that it cannot be that $2 < y < 3$, and hence $(y,z)$ cannot be in $S_5$.
    \end{indpar}

    Thus in every case that is actually possible we have that $(x,z) \in T$, which shows that $T$ is transitive.

    Therefore $T$ is an equivalence relation that contains $S$ so that $T \sps \bigcap_{R \in S^*} R$ as discussed above, which of course completes the proof that $T = \bigcap_{R \in S^*} R$.
  }

  As far as the equivalence classes formed by $T$ are concerned, refer to the illustration above.
  Consider the equivalence class $C$ contains $x \in \reals$.
  If $x \leq 0$ or $x \geq 3$ then $C = \braces{x}$ because there is no other $y$ for which $yTx$ except $y=x$.
  So suppose that $0 < x < 3$.
  If $x$ is an integer so that $x = 1$ or $x = 2$, then $C = \braces{1,2}$.
  If $x$ is not an integer, then $C$ always has three elements.
  We have that
  \gath{
    C = \begin{cases}
      \braces{x, x+1, x+2} & 0 < x < 1 \\
      \braces{x-1, x, x+1} & 1 < x < 2 \\
      \braces{x-2, x-1, x} & 2 < x < 3
    \end{cases} \,.
  }
  These facts can be deduced by examining where the vertical line intersecting the $x$-axis at $x$ intersects the graph of $T$.
}

\exercise{6}{
  Define a relation on the plane by setting
  \gath{
    \ppt{0} < \ppt{1}
  }
  if either $y_0 - x_0^2 < y_1 - x_1^2$ or $y_0 - x_0^2 = y_1 - x_1^2$ and $x_0 < x_1$.
  Show that this is an order relation on the plane, and describe it geometrically.
}
\sol{
  First we show that $<$ is an order relation on the plane.
  \qproof{
    As clearly $<$ is a relation on the plane, we need only show that it has the three required properties of an order relation:

    (Comparability) Consider distinct $\ppt{0}$ and $\ppt{1}$ in the plane so that either $x_0 \neq x_1$ or $y_0 \neq y_1$.
    Obviously if $\peq{0} < \peq{1}$ (or $\peq{1} < \peq{0}$) then of course $\ppt{0} < \ppt{1}$ (or $\ppt{1} < \ppt{0}$) so we are done.
    So assume that $\peq{0} = \peq{1}$.
    It it were the case that $x_0 = x_1$ it would have to be that $y_0 \neq y_1$, but we would have
    \gath{
      \peq{0} = \peq{1} = y_1 - x_0^2 \\
      y_0 = y_1 \,,
    }
    which is a contradiction.
    So it must be that $x_0 \neq x_1$.
    So either $x_0 < x_1$ and so $\ppt{0} < \ppt{0}$ or $x_1 < x_0$ and so $\ppt{1} < \ppt{0}$.
    This shows that $<$ is comparable in the plane.

    (Nonreflexivity) Consider $\ppt{}$ in the plane so that obviously $\peq{} = \peq{}$.
    As we also have that $x = x$, it is not the case that $x < x$ so that it is not true that $\ppt{} < \ppt{}$.

    (Transitivity) Suppose that $\ppt{0} < \ppt{1}$ and $\ppt{1} < \ppt{2}$.
    We then have the following:

    Case: $\peq{0} < \peq{1}$.
    \begin{indpar}
      Case:  $\peq{1} < \peq{2}$.
      Then of course $\peq{0} < \peq{1} < \peq{2}$ so that $\ppt{0} < \ppt{2}$.

      Case: $\peq{1} = \peq{2}$ and $x_1 < x_2$.
      Then we have $\peq{0} < \peq{1} = \peq{2}$ so that again $\ppt{0} < \ppt{2}$.
    \end{indpar}

    Case: $\peq{0} = \peq{1}$ and $x_0 < x_1$.
    \begin{indpar}
      Case: $\peq{1} < \peq{2}$.
      Then $\peq{0} = \peq{1} < \peq{2}$ so that $\ppt{0} < \ppt{2}$.

      Case: $\peq{1} = \peq{2}$ and $x_1 < x_2$.
      Then $\peq{0} = \peq{1} = \peq{2}$ and $x_0 < x_1 < x_2$ so that $\ppt{0} < \ppt{2}$.
    \end{indpar}

    Thus in all cases $\ppt{0} < \ppt{2}$, which shows that $<$ is transitive in the plane.
  }

  Geometrically, we refer back to Exercise~3.1 and consider a parabola $y = x^2$ shifted up or down on the $y$-axis be some amount $c$.
  Then $y = x^2 + c$ so that $\peq{} = c$, and hence every $\ppt{}$ point on the parabola has the same value for $\peq{}$, namely $c$.
  Therefore if two distinct points $\ppt{0}$ and $\ppt{1}$ lie on the different parabolas then $\peq{0}$ and $\peq{1}$ will have different values, say $c$ and $d$, respectively.
  Then clearly if $\ppt{1}$ is on a higher parabola on the $y$-axis then $c < d$ so that $\peq{0} = c < d = \peq{1}$ so that $\ppt{0} < \ppt{1}$ in our order.
  If the points lie on the same parabola then $\peq{0} = \peq{1}$ and whichever points is further to the right will be larger in our order since then, for example, $x_0 < x_1$ so that $\ppt{0} < \ppt{1}$.
}

\exercise{7}{
  Show that the restriction of an order relation is an order relation.
}
\sol{
  \qproof{
    Suppose that $A$ is a set with order relation $<$.
    Also let $A_0$ be a subset of $A$ so that $\prec = < \cap (A_0 \times A_0)$ is the restriction of $<$ to $A_0$.
    Clearly $\prec \ss A_0 \times A_0$ so that it s a relation on $A_0$.
    So we must show that $\prec$ satisfies the three properties of an order relation:

    (Comparability) Consider any $x,y \in A_0$ so that also $x,y \in A$ since $A_0 \ss A$.
    Then $x$ and $y$ are comparable in $<$ since it is an order.
    So, without loss of generality, we can assume that $x < y$ and so $(x,y) \in <$.
    Since clearly also $(x,y) \in A_0 \times A_0$, it follows that $(x,y) \in < \cap (A_0 \times A_0) = \prec$ and hence $x \prec y$.
    This shows that $x$ and $y$ are comparable in $\prec$.

    (Nonreflexivity) Suppose that $x \in A_0$ so that also $x \in A$ since $A_0 \ss A$.
    Then it cannot be true that $x < x$ since $A$ is an order and so is nonreflexive.
    Thus $(x,x) \notin <$ so that also $(x,x) \notin < \cap (A_0 \times A_0) = \prec$.
    Hence it is not true that $x \prec x$ so that $\prec$ is also nonreflexive since $x$ was arbitrary.

    (Transitivity) Suppose that $x \prec y$ and $y \prec z$.
    Then of course $(x,y) \in A_0 \times A_0$ and $(x,y) \in <$, and similarly $(y,z) \in A_0 \times A_0$ and $(y,z) \in <$.
    Hence $x < y$ and $y < z$ so that $x < z$ since $<$ is an order and therefore transitive.
    Thus $(x,z) \in <$ so that $(x,z) \in < \cap (A_0 \times A_0) = \prec$ since clearly $(x,z) \in A_0 \times A_0$.
    So then $x \prec z$, which shows that $\prec$ is transitive.
  }
}

\exercise{8}{
  Check that the relation defined in Example~7 is an order relation.
}
\sol{
  Recall that Example~7 is the relation on $C$ on the real line such that $xCy$ if $x^2 < y^2$ or $x^2 = y^2$ and $x < y$.
  \qproof{
    We show that this satisfies the three properties of an order:

    (Comparability) Suppose that $x$ and $y$ are distinct real numbers.
    If $x^2 < y^2$ (or $y^2 < x^2$) then of course $xCy$ (or $yCx$) so we are done.
    So assume that $x^2 = y^2$.
    Since we know that $x \neq y$, it has to be that $y = -x$ so that still $x^2 = y^2$.
    This also implies that $x,y \neq 0$ since otherwise we would have $0 = y = -x = 0 = x$.
    If $x > 0$ then we have $y = -x < 0 < x$.
    If $x < 0$ then we have $x < 0 < -x = y$.
    Hence either way $x^2 = y^2$ but $x < y$ (or $y < x$) so that $xC$ (or $yCx$).
    This shows that $x$ and $y$ are comparable in $C$.

    (Nonreflexivity) Suppose that $x \in \reals$ so that of course $x^2 = x^2$.
    However clearly it is not the case that $x < x$ so that it cannot be that $xCx$ in this relation.

    (Transitivity) Suppose that $xCy$ and $yCz$.
    We then have the following cases:

    Case: $x^2 < y^2$.
    \begin{indpar}
      Case: $y^2 < z^2$.
      Then clearly $x^2 < y^2 < z^2$ so that $xCz$.

      Case: $y^2 = z^2$ and $y < z$.
      Then $x^2 < y^2 = z^2$ so that again $xCz$.
    \end{indpar}

    Case: $x^2 = y^2$ and $x < y$.
    \begin{indpar}
      Case: $y^2 < z^2$.
      Then we have $x^2 = y^2 < z^2$ so that $xCz$.
      
      Case: $y^2 = z^2$ and $y < z$.
      Then $x^2 = y^ = z^2$ and $x < y < z$ so that again $xCz$.
    \end{indpar}

    Hence in all cases $xCz$ so that $C$ is also transitive.
  }

  We note that this order relation differs from the normal order on $\reals$.
  For example if $x = -2$ and $y = 1$ then clearly $x < y$ in the normal order.
  However, we have that $y^2 = 1^2 = 1 < 4 = (-2)^2 = x^2$ so that $yCx$.
}

\newcommand\appt[1]{a_{#1} \times b_{#1}}
\exercise{9}{
  Check that the dictionary order is an order relation.
}
\sol{
  Suppose that $A$ and $B$ are two sets with order relations $<_A$ and $<_B$.
  Recall that the dictionary order $<$ on $A \times B$ is defined by
  \gath{
    \appt{1} < \appt{2}
  }
  if $a_1 <_A a_2$ or if $a_1 = a_2$ and $b_1 <_B b_2$.
  \qproof{
    Clearly $<$ is a relation on $A \times B$ so we just need to show the three properties of an order relation:

    (Comparability) Consider distinct points $\appt{1}$ and $\appt{2}$ in $A \times B$ so that $a_1 \neq a_2$ or $b_1 \neq b_1$.
    If $a_1 \neq a_2$ then they are comparable in $<_A$ (since it is an order relation) so that, without loss of generality, we can assume that $a_1 <_A a_2$.
    Then of course $\appt{1} < \appt{2}$ by definition.
    So assume that $a_1 = a_2$ so that it must be that $b_1 \neq b_2$.
    Then $b_1$ and $b_2$ are comparable in $<_B$ since it is an order relation.
    So without loss of generality we can assume that $b_1 <_B b_2$.
    Then of course $\appt{1} < \appt{2}$ since also $a_1 = a_2$.
    Thus either way $\appt{1}$ and $\appt{2}$ are comparable in $<$.

    (Nonreflexivity) Suppose that $\appt{}$ is any element of $A \times B$.
    Since $<_B$ is an order relation, it is nonreflexive so that it is not true that $b <_B b$.
    Since of course $a=a$, it follows that it cannot be that $\appt{} < \appt{}$ since it would have to be that $b <_B b$.
    Hence $<$ is nonreflexive since $\appt{}$ was arbitrary.

    (Transitivity) Suppose that $\appt{1} < \appt{2}$ and $\appt{2} < \appt{3}$.
    We then have the following cases:

    Case: $a_1 <_A a_2$.
    \begin{indpar}
      Case: $a_2 <_A a_3$.
      Then $a_1 <_A a_2$ and $a_2 <_A a_3$ so that $a_1 <_A a_3$ since $<_A$ is transitive.
      Thus by definition $\appt{1} < \appt{3}$.

      Case: $a_2 = a_3$ and $b_2 <_B b_3$.
      Then $a_1 <_A a_2 = a_3$ so that $\appt{1} < \appt{3}$ by definition.
    \end{indpar}

    Case: $a_1 = a_2$ and $b_1 <_B b_2$.
    \begin{indpar}
      Case: $a_2 <_A a_3$.
      Then $a_1 = a_2 <_A a_3$ so that by definition $\appt{1} < \appt{3}$.

      Case: $a_2 = a_3$ and $b_2 <_B b_3$.
      Then $a_1 = a_2 = a_3$ and $b_1 <_B b_2$ and $b_2 <_B b_3$ so that $b_1 <_B b_3$ since $<_B$ is transitive.
      Therefore again $\appt{1} < \appt{3}$ by definition.
    \end{indpar}

    Thus in all cases $\appt{1} < \appt{3}$, which shows that $<$ is transitive.
  }
}

\def\srt{\sqrt{1+4y^2}}
\exercise{10}{
  \eparts{
  \item Show that the map $f: (-1,1) \to \reals$ of Example~9 is order preserving.
  \item Show that the equation $g(y) = 2y/[1+(1+4y^2)^{1/2}]$ defines a function $g: \reals \to (-1,1)$ that is both a left and right inverse for $f$.
  }
}
\sol{
  (a)
  Recall that $f$ from Example~3.9 is defined by
  \gath{
    f(x) = \frac{x}{1-x^2}
  }
  for $x \in (-1,1)$.
  We show that $f$ preserves order.
  \qproof{
    Suppose that $x,y \in (-1,1)$ and that $x < y$.
    Now, since $-1 < x < 1$, we have $0 \leq \abs{x} < 1$ so that
    \gath{
      x^2 = \abs{x}^2 = \abs{x}\abs{x} < \abs{x} \cdot 1 = \abs{x} < 1.
    }
    Hence $0 < 1-x^2$ so that also $0 < 1/(1-x^2)$ as well.
    By the same argument $0 < 1-y^2$ so that $0 < 1/(1-y^2)$.
    We then have the following cases:

    Case: $x \geq 0$.
    Then also $y > 0$ since $y > x \geq 0$.
    Thus
    \ali{
      x^2 = x \cdot x \leq x &\cdot y < y \cdot y = y^2 & \text{(since $x<y$, $x \geq 0$, and $y > 0$)} \\
      -x^2 &> -y^2 \\
      1-x^2 &> 1-y^2 \\
      \frac{1}{1-y^2} &> \frac{1}{1-x^2} \,.
    }
    From this it follows that
    \gath{
      f(x) = \frac{x}{1-x^2} < \frac{y}{1-x^2} < \frac{y}{1-y^2} = f(y)
    }
    since $x < y$ and $1/(1-x^2)$ and $1/(1-y^2)$ are both positive.

    Case: $x < 0$.
    Then $f(x) = x/(1-x^2) < 0$ since $x<0$ and $1/(1-x^2) > 0$.
    \begin{indpar}
      Case: $y \leq 0$.
      Then we have that $-x > -y \geq 0$ so that $f(-y) < f(-x)$ by the previous case.
      But we have $f(-x) = -x/(1-(-x)^2) = -x/(1-x^2) = -f(x)$, and similarly $f(-y) = -f(y)$.
      Hence $-f(y) = f(-y) < f(-x) = -f(x)$ so that $f(y) > f(x)$.

      Case: $y > 0$.
      Then $f(y) = y/(1-y^2) > 0$ since $y>0$ and $1/(1-y^2) > 0$.
      Hence $f(x) < 0 < f(y)$.
    \end{indpar}

    Therefore in all cases we have $f(x) < f(y)$, which shows that $f$ preserves order.
  }

  (b)
  \qproof{
    First we need to show that $g$ is even a function from $\reals$ to $(-1,1)$.
    We note that
    \gath{
      y^2 \geq 0 \\
      4y^2 \geq 0 \\
      1+4y^2 \geq 1 > 0 \\
      \srt > 0 \\
      1 + \srt > 1 > 0
    }
    so that $g(y)$ is well-defined for all $y \in \reals$ since the denominator of $g(y)$ is always nonzero.
    Now, if $y = 0$ then clearly $\abs{g(y)} = \abs{g(0)} = \abs{0} = 0 < 1$.
    So in what follows assume that $y \neq 0$ so that $\abs{y} > 0$ and $y^2 > 0$.
    Then we have
    \gath{
      1 > 0 \\
      1 + 4y^2 > 4y^2 > 0 \\
      \srt > \sqrt{4y^2} = 2\abs{y} \\
      1 + \srt > 1 + 2\abs{y} > 2 \abs{y} > 0 \\
      \frac{1}{2\abs{y}} > \frac{1}{1+\srt} \\
      1 = \frac{2\abs{y}}{2\abs{y}} > \frac{2\abs{y}}{1+\srt} = \abs{\frac{2y}{1+\srt}} = \abs{g(y)}
    }
    so that $-1 < g(y) < 1$ and hence $g(y) \in (-1,1)$.
    Thus $g$ is in fact a well-defined function from $\reals$ into $(-1,1)$.

    To show that $g$ is a left inverse of $f$ consider any $x \in (-1,1)$.
    First we have
    \ali{
      \sqrt{1 + 4f(x)^2} &= \sqrt{1 + \frac{4x^2}{(1-x^2)^2}} = \sqrt{\frac{(1-x^2)^2 + 4x^2}{(1-x^2)^2}} \\
      &= \sqrt{\frac{1 - 2x^2 + x^4 + 4x^2}{{(1-x^2)^2}}} = \sqrt{\frac{1 + 2x^2 + x^4}{(1-x^2)^2}} \\
      &= \sqrt{\frac{(1 + x^2)^2}{(1-x^2)^2}} = \frac{1+x^2}{1-x^2}
    }
    so that
    \ali{
      (g \circ f)(x) &= g(f(x)) = \frac{2f(x)}{1 + \sqrt{1+4f(x)^2}} = \frac{2x/(1-x^2)}{1 + \frac{1+x^2}{1-x^2}} \\
      &= \frac{2x}{(1-x^2)\parens{1 + \frac{1+x^2}{1-x^2}}} = \frac{2x}{1 - x^2 + 1 + x^2} \\
      &= \frac{2x}{2} = x \,,
    }
    which shows that $g \circ f = i_{(-1,1)}$ since $x$ was arbitrary so that $g$ is a left inverse of $f$.

    Now consider any $y \in \reals$.
    Then we first have
    \ali{
      g(y)^2 &= \parens{\frac{2y}{1 + \srt}}^2 = \frac{4y^2}{(1 + \srt)^2} \\
      &= \frac{4y^2}{1 + 2\srt + 1 + 4y^2} = \frac{4y^2}{2 + 4y^2 + 2\srt} \\
      &= \frac{2y^2}{1 + 2y^2 + \srt}
    }
    so that
    \ali{
      (f \circ g)(y) &= f(g(y)) = \frac{g(y)}{1 - g(y)^2} = \frac{2y/(1 + \srt)}{1 - \frac{2y^2}{1 + 2y^2 + \srt}} \\
      &= \frac{2y}{(1 + \srt)\parens{\frac{1 + 2y^2 + \srt - 2y^2}{1 + 2y^2 + \srt}}} \\
      &= \frac{2y}{(1 + \srt)\parens{\frac{1 + \srt}{1 + 2y^2 + \srt}}} \\
      &= \frac{2y}{\frac{1 + 2\srt + 1 + 4y^2}{1 + 2y^2 + \srt}} = \frac{2y}{\frac{2 + 4y^2 + 2\srt}{1 + 2y^2 + \srt}} \\
      &= \frac{2y}{2\parens{\frac{1 + 2y^2 + \srt}{1 + 2y^2 + \srt}}} = \frac{2y}{2} = y \,,
    }
    which shows that $f \circ g = i_\reals$ since $y$ was arbitrary so that $g$ is also a right inverse of $f$.
  }
  Note that what was shown implies that $f$ is bijective and that $g$ is its inverse, by Exercise~2.5 part~(e).
}

\exercise{11}{
  Show that an element in an ordered set has at most one immediate successor and at most one immediate predecessor.
  Show that a subset of an ordered set has at most one smallest element and at most one largest element.
}
\sol{
  \begin{lem}\label{lem:relations:lngeq}
    Suppose that $A$ is a set with order $<$ and that $a$ and $b$ are two elements of $A$.
    Then $a < b$ if and only if it is not true that $b \leq a$.
  \end{lem}
  \qproof{
    $(\imp)$ Suppose that $a < b$.
    If it were the case that $a = b$ then we would have $a < b = a$, which would violate the nonreflexivity property of the order.
    If it were the case that $b < a$ then we would have $a < b$ and $b < a$ so that $a < a$ by the transitive property of the order.
    This again violates nonreflexivity.
    Hence neither $a=b$ nor $b < a$ so that it is not true that $b \leq a$.

    ($\pmi$) Now suppose that it is \emph{not} true that $b \leq a$.
    Then neither $b = a$ nor $b < a$.
    Since $b \neq a$, it must be that either $a < b$ or $b < a$ by the comparability property.
    However, we know that cannot be that $b < a$, so it must be that $a < b$.
  }
  
  \mainprob
  
  \qproof{
    In what follows Suppose that $A$ is a set with order $<$.

    First let $a$ be an element of $A$ and suppose that $b_1$ and $b_2$ are both immediate successors of $a$, which of course means that $a < b_1,b_2$.
    Then by definition the open intervals  $(a,b_1)$ and $(a,b_2)$ are both empty.
    Now, suppose that $b_1$ and $b_2$ are distinct so that they must be comparable since $<$ is an order.
    Without loss of generality we can assume that $b_1 < b_2$, but then we have $a < b_1 < b_2$ so that $b_1 \in (a,b_2)$.
    This is a contradiction since we know that $(a,b_2)$ is empty, so it has to be that $b_1 = b_2$.
    This of course shows that the immediate successor is unique.
    An analogous argument shows that the immediate predecessor, if it exists, is also unique.

    Now suppose that $A_0$ is a subset of $A$ with smallest elements $a_1$ and $a_2$.
    If $a_1$ and $a_2$ were to be distinct then they must be comparable so that we can assume $a_1 < a_2$.
    However, then it is not true that $a_2 \leq a_1$ by Lemma~\ref{lem:relations:lngeq}, but this means that $a_2$ cannot be a smallest element of $A_0$ since $a_1 \in A_0$.
    As this is a contradiction, it must be that $a_1 = a_2$, which shows that the smallest element is unique if there is one.
    An analogous argument shows any largest element is also unique.
  }
}

\def\pintss{\pints \times \pints}
\exercise{12}{
  Let $\pints$ denote the set of positive integers.
  Consider the following order relations on $\pintss$.
  \begin{enumerate}[label=(\roman*)]
  \item The dictionary order.
  \item $\ppt{0} < \ppt{1}$ if either $x_0 - y_0 < x_1 - y_1$ or $x_0 - y_0 = x_1 - y_1$ and $y_0 < y_1$.
  \item $\ppt{0} < \ppt{1}$ if either $x_0 + y_0 < x_1 + y_1$ or $x_0 + y_0 = x_1 + y_1$ and $y_0 < y_1$.
  \end{enumerate}
  In these order relations, which elements have immediate predecessors?
  Does the set have a smallest element?
  Show that the three order types are different.
}
\newcommand\draworder[1]{
  % Coordinate grid
  \draw[step=1cm,\tikzgridclr,thin] (0,0) grid (4.9,4.9);

  % Axes
  \draw[thick,->] (0,0) -- (5,0);
  \draw[thick,->] (0,0) -- (0,5);

  #1

  % Points
  \foreach \x in {1,2,3,4}
  \foreach \y in {1,2,3,4}
  \filldraw[fill=\tikzgray] (\x,\y) circle (0.1);
}
\sol{
  \begin{lem}\label{lem:relations:otsmallest}
    If $A$ and $B$ are ordered sets and $A$ has a smallest element but $B$ does not, then $A$ and $B$ do not have the same order type.
  \end{lem}
  \qproof{
    This may seem fairly obvious but we show it formally anyway.
    Suppose that $<_A$ and $<_B$ are the orders on $A$ and $B$, respectively.
    Let $a$ be the smallest element of $A$ and suppose to the contrary that they \emph{do} have the same order type.
    Then there is a bijection $f: A \to B$ that preserves order, noting that $\inv{f}$ is also then a bijection.
    Now, since $B$ has no smallest element, there must be a $b \in B$ where $b <_B f(a)$.
    Setting $a' = \inv{f}(b)$ so that $f(a') = b$ we have that $f(a') = b <_B f(a)$, and hence it has to be that $a' <_A a$ since otherwise we'd have $a \leq_A a'$ so that $f(a) \leq_B f(a')$.
    However, $a' < a$ means that it is not true that $a \leq a'$ by Lemma~\ref{lem:relations:lngeq}, which contradicts the fact that $a$ is the smallest element of $A$.
    Therefore it must be that no such $f$ exists so that $A$ and $B$ have different order types.
    
  }

  \mainprob
  
  The figure below illustrates the dictionary order of part~(i):
  \begin{center}
    \begin{tikzpicture}[scale=0.8]
      \draworder{
        % Order lines
        \foreach \x in {1,2,3,4}
        \draw[thick,dashed,->] (\x,1) -- (\x,4.9);
        \foreach \x in {1,2,3}
        \draw[thick,dashed,->] (\x,4.9) -- (\x+0.9,1+0.1);
      }
    \end{tikzpicture}
  \end{center}
  We claim that every point has an immediate predecessor except points in the subset
  \gath{
    A = \braces{\ppt{} \where y = 1} \,.
  }
  We also claim that $(1,1)$ is the smallest element in $\pintss$ with this order.
  \qproof{
    First consider any point $\ppt{1} \in \pintss$ where $\ppt{1} \notin A$ so that $y_0 \neq 1$.
    It then follows that $y_1 > 1$ since $1$ is the smallest positive integer.
    Then set $x_0 = x_1$ and $y_0 = y_1 - 1$ so that clearly $\ppt{0} \in \pintss$ since $y_0 > 0$ because $y_1 > 1$.
    Clearly also $x_0 = x_1$ and $y_0 < y_1$ so that $\ppt{0} < \ppt{1}$ in the dictionary order.
    We claim that $\ppt{0}$ is the immediate predecessor of $\ppt{1}$.
    So suppose to the contrary that there is an $\ppt{2} \in \pintss$ such that $\ppt{0} < \ppt{2} < \ppt{1}$.
    It cannot be that $x_0 < x_2$ since then we would have $x_1 = x_0 < x_2$ so that $\ppt{1} < \ppt{2}$, which by Lemma~\ref{lem:relations:lngeq} contradicts the fact that $\ppt{2} < \ppt{1}$.
    So it has to be that $x_0 = x_2$ and $y_0 < y_2$.
    Since then $x_2 = x_0 = x_1$, it must also be that $y_2 < y_1$ since $\ppt{2} < \ppt{1}$.
    But then we have $y_0 < y_2 < y_1 = y_0+1$, which is not possible since $y_1 = y_0+1$ is the immediate successor of $y_0$ in $\pints$ so that there can be no integers between them.
    So it must be that no $\ppt{2}$ exists so that $\ppt{0}$ is the immediate predecessor of $\ppt{1}$.
    Thus every element not in $A$ has an immediate predecessor.

    Now consider any $\ppt{1} \in A$ so that $y_1 = 1$, and consider any point $\ppt{0} < \ppt{1}$.
    It cannot be that $x_0 = x_1$ since then we would have $y_0 < y_1 = 1$, which is not possible since $1$ is the smallest positive integer.
    So it must be that $x_0 < x_1$.
    Then let $x_2 = x_0$ and $y_2 = y_0+1$ so that clearly $\ppt{2} \in \pintss$ since $y+1$ is always still a positive integer if $y$ is.
    Then we have that $x_0 = x_2$ and $y_0 < y_0 + 1 = y_2$, and hence $\ppt{0} < \ppt{2}$.
    We also have $x_2 = x_0 < x_1$ so that $\ppt{2} < \ppt{1}$.
    Therefore $\ppt{0} < \ppt{2} < \ppt{1}$, which shows that $\ppt{0}$ is not the immediate predecessor of $\ppt{1}$.
    This shows that $\ppt{1}$ has no immediate predecessor since $\ppt{0} < \ppt{1}$ was arbitrary.
    Since $\ppt{1} \in A$ was arbitrary, this completes the proof that every element in $\pintss$ has an immediate predecessor except those in $A$.

    It is easy to prove that $(1,1)$ is the smallest element in the dictionary order.
    Consider any $\ppt{0} \in \pintss$ and suppose that $\ppt{0} < (1,1)$.
    It cannot be that $x_0 < 1$ since $1$ is the smallest positive integer.
    So then $x_0 = 1$ and $y_0 < 1$, but this is also not possible, again since $1$ is the smallest positive integer.
    Thus it cannot be that $\ppt{0} < (1,1)$, so it must be that $(1,1) \leq \ppt{0}$ by Lemma~\ref{lem:relations:lngeq}.
    This shows that $(1,1)$ is the smallest element since $\ppt{0}$ was arbitrary.
  }

  Below is shown an illustration for the order in part~(ii):
  \begin{center}
    \begin{tikzpicture}[scale=0.8]
      \draworder{
        % Order lines
        \foreach \y in {1,2,3,4}
        \draw[thick,dashed,->] (1,\y) -- (5.9-\y,4.9);
        \foreach \x in {2,3,4}
        \draw[thick,dashed,->] (\x,1) -- (4.9,5.9-\x);
        \foreach \y in {2,3,4}
        \draw[thick,dashed,->] (5.9-\y,4.9) -- (1+0.1,\y-1+0.1);
        \draw[thick,dashed,->] (1.3,4.9) -- (1+0.1,4+0.1);
        \foreach \x in {1,2,3}
        \draw[thick,dashed,->] (4.9,5.9-\x) -- (\x+1+0.1,1+0.1);
      }
    \end{tikzpicture}
  \end{center}
  We claim that every element has an immediate predecessor except those in the subset
  \gath{
    A = \braces{\ppt{} \where x=1 \text{ or } y=1} \,.
  }
  We also claim that the set $\pintss$ has no smallest element in this order.
  \qproof{
    First consider any $\ppt{1} \notin A$ so that $x_1 \neq 1$ and $y_1 \neq 1$.
    Then it has to be that $x_1,y_1 > 1$.
    So set $\ppt{0} = (x_1-1,y_1-1)$ so that clearly still $\ppt{0} \in \pintss$.
    Then $x_0-y_0 = (x_1-1)-(y_1-1) = x_1-1-y_1+1 = x_1-y_1$.
    We also have $y_0 = y_1-1 < y_1$ so that $\ppt{0} < \ppt{1}$.
    Now suppose that there is an $\ppt{2} \in \pintss$ where $\ppt{0} < \ppt{2} < \ppt{1}$.
    It cannot be that $x_0-y_0 < x_2-y_2$ since then we would have $x_1-y_1 = x_0-y_0 < x_2-y_2$ so that $\ppt{1} < \ppt{2}$, which we know cannot be the case since $\ppt{2} < \ppt{1}$.
    So it must be that $x_2-y_2 = x_0-y_0 = x_1-y_1$ and $y_0 < y_2$, but then we must have $y_0 < y_2 < y_1 = y_0+1$, noting that $y_2 < y_1$ because $x_2-y_2 = x_1-y_1$ and $\ppt{2} < \ppt{1}$.
    However, this is not possible since of course $y_0+1$ is the immediate successor of $y_0$ in $\pints$.
    So then it has to be that no such $\ppt{2}$ exists so that $\ppt{0}$ is the immediate predecessor of $\ppt{1}$.
    This shows that every point not in $A$ has an immediate predecessor since $\ppt{1}$ was arbitrary.

    Now suppose that $\ppt{1} \in A$ so that $x_1 = 1$ or $y_1 = 1$, and consider any $\ppt{0} < \ppt{1}$.

    Case: $x_1 = 1$.
    Suppose that $x_0-y_0 = x_1-y_1$ and $y_0 < y_1$.
    Then we would have $-y_0 > -y_1$ and $x_0-y_0 = x_1-y_1 = 0 - y_1 = -y_1 < -y_0$ so that $x_0 < 0$ (by adding $y_0$ to both sides), which is not possible.

    Case: $y_1 = 1$.
    It clearly cannot be that case that $x_0-y_0 = x_1-y_1$ and $y_0 < y_1$ since then $y_0 < y_1 = 1$, which is not possible.

    So in either case it must be that $x_0-y_0 < x_1-y_1$ since $\ppt{0} < \ppt{1}$.
    So set the point $\ppt{2} = (x_0+1,y_0+1)$, which is clearly still an element of $\pintss$.
    We then have $x_2-y_2 = (x_0+1)-(y_0+1) = x_0+1-y_0-1 = x_0-y_0 < x_1-y_1$ so that $\ppt{2} < \ppt{1}$.
    We also have $x_0-y_0 = x_2-y_2$ and $y_0 < y_0+1 = y_2$ so that also $\ppt{0} < \ppt{2}$.
    Hence $\ppt{0} < \ppt{2} < \ppt{1}$ so that $\ppt{0}$ is not the immediate predecessor of $\ppt{1}$.
    Since $\ppt{0} < \ppt{1}$ was arbitrary, this shows that $\ppt{1}$ has no immediate predecessor at all.
    Since $\ppt{1} \in A$ was arbitrary, this shows that no element of $A$ has an immediate predecessor.

    To show that $\pints$ in this order has no smallest element, consider absolutely any $\ppt{1} \in \pintss$.
    Let $\ppt{0} = (x_1,y_1+1)$ so that of course $\ppt{0} \in \pintss$.
    We then have that $x_0 - y_0 = x_1 - (y_1+1) = (x_1 - y_1) - 1 < x_1 - y_1$ so that $\ppt{0} < \ppt{1}$.
    Then of course is it not true that $\ppt{1} \leq \ppt{0}$ by Lemma~\ref{lem:relations:lngeq} so that $\ppt{1}$ cannot be the smallest element.
    Then, since $\ppt{1}$ was arbitrary, this shows that $\pintss$ has no smallest element in this order.
  }

  An illustration of the order of part~(iii) is shown below:
  \begin{center}
    \begin{tikzpicture}[scale=0.8]
      \draworder{
        % Order lines
        \foreach \x in {2,3,4}
        \draw[thick,dashed,->] (\x,1) -- (1+0.1,\x-0.1);
        \foreach \x in {5,6,7}
        \draw[thick,dashed,->] (4.9,\x-3.9) -- (\x-3.9,4.9);
        \draw[thick,dashed,->] (1,1) -- (2-0.1,1);
        \foreach \y in {2,3,4}
        \draw[thick,dashed,->] (1,\y) -- (\y+1-0.1,1+0.1);
        \clip (0,0) rectangle (4.9,4.9);
        \foreach \y in {5,6,7}
        \draw[thick,dashed,->] (1,\y) -- (4.9,\y-3.2);
      }
    \end{tikzpicture}
  \end{center}
  We claim that every element has an immediate predecessor except for $(1,1)$, which is the smallest element.
  \qproof{
    First we show that $\ppt{0} = (1,1)$ is the smallest element, from which it follows that it cannot have an immediate predecessor since it has no predecessors at all.
    Consider any $\ppt{1}$ in $\pintss$.
    If $\ppt{1} = (1,1)$ then of course $\ppt{0} = (1,1) \leq \ppt{1}$ is true, so assume that $\ppt{1} \neq (1,1)$ so that either $x_1 \neq 1$ or $y_1 \neq 1$.
    If $x_1 \neq 1$ then it has to be that $x_1 > 1$ so that $x_0 + y_0 = 1 + 1 \leq 1 + y_1 < x_1 + y_1$, and hence $\ppt{0} < \ppt{1}$.
    If $y_1 \neq 1$ then $y_1 > 1$ so that again $x_0 + y_0 = 1 + 1 \leq x_1 + 1 < x_1 + y_1$, and hence $\ppt{0} < \ppt{1}$.
    Thus in every case $\ppt{0} \leq \ppt{1}$, which shows that $\ppt{0} = (1,1)$ is the smallest element since $\ppt{1}$ was arbitrary.

    Now we show that every other element of $\pints$ has an immediate predecessor in this order.
    So consider any $\ppt{1} \in \pintss$ where $\ppt{1} \neq (1,1)$.
    Hence either $x_1 \neq 1$ or $y_1 \neq 1$.

    Case: $y_1 = 1$.
    Then it has to be that $x_1 \neq 1$ so that $x_1 > 1$.
    We claim that $\ppt{0} = (1, x_1-1)$ is the immediate predecessor of $\ppt{1}$.
    First we note that clearly $\ppt{0} \in \pintss$ since $x_1 > 1$.
    We also have that $x_0 + y_0 = 1 + x_1 - 1 = x_1 < x_1 + y_1$ since $0 < 1 = y_1$, and so $\ppt{0} < \ppt{1}$.
    Now suppose that there is a point $\ppt{2} \in \pintss$ where $\ppt{0} < \ppt{2} < \ppt{1}$.
    It cannot be that $x_1+y_1 = x_2+y_2$ and $y_1 < y_2$ because then we would have $y_2 < y_1 = 1$, which is not possible since $y_2 \in \pints$.
    So it must be that $x_2+y_2 < x_1+y_1 = x_1+1$ since $\ppt{2} < \ppt{1}$.
    Now, since also $\ppt{0} < \ppt{2}$, it must be that $x_0+y_0 \leq x_2+y_2$, but then we have $x_1 = 1 + (x_1 - 1) = x_0 + y_0 \leq x_2+y_2 < x_1+1$, which is impossible.
    So it has to be that no such $\ppt{2}$ exists so that $\ppt{0}$ is the immediate predecessor of $\ppt{1}$.

    Case: $y_1 \neq 1$.
    Then it has to be that $y_1 > 1$ so that the point $\ppt{0} = (x_1+1,y_1-1)$ is still an element of $\pintss$.
    We show that $\ppt{0}$ is the immediate predecessor of $\ppt{1}$.
    First we have that $x_0+y_0 = (x_1+1) + (y_1-1) = x_1 + y_1$ and $y_0 = y_1-1 < y_1$ so that $\ppt{0} < \ppt{1}$.
    Now suppose that there a point $\ppt{2} \in \pintss$ where $\ppt{0} < \ppt{2} < \ppt{1}$.
    Then it has to be that $x_0 + y_0 \leq x_2+y_2 \leq x_1+y_1$ so that we have $x_1+y_1 = (x_1+1)+(y_1-1) = x_0+y_0 \leq x_2+y_2 \leq x_1+y_1$ so that $x_0+y_0 = x_2+y_2 = x_1+y_1$.
    But then we must have $y_1 - 1 = y_0 < y_2 < y_1$, which is not possible since $y_1-1$ is the immediate predecessor of $y_1$ in $\pints$.
    So no such $\ppt{2}$ can exists, and hence $\ppt{0}$ is the immediate predecessor of $\ppt{1}$.

    This in all cases $\ppt{1}$ has an immediate predecessor, which shows the desired result since $\ppt{1} \neq (1,1)$ was arbitrary.
  }

  Now we show that all three orders have different order types.
  \qproof{
    It follows immediately from Lemma~\ref{lem:relations:otsmallest} that order (i) and order (ii) do not have the same order type since (i) has a smallest element while (ii) does not.
    Similarly order (iii) and order (ii) cannot have the same order type for the same reason.
    So all that remains to be shown is that orders (i) and (iii) have different order types.

    So denote order (i) with $<$ and order (iii) with $\prec$ and suppose to the contrary that they do have the same order type.
    Then there is a bijection $f: \pintss \to \pintss$ that preserves order, supposing that the domain has the dictionary order $<$ and the range has the order $\prec$.
    Then of course $\inv{f}$ is also a bijection that preserves order.
    It was shown above that $\pintss$ with $<$ has countably many elements with no immediate predecessor, whereas $\pintss$ with $\prec$ has only a single such element, namely the smallest element $(1,1)$.

    Thus we can choose an element $\ppt{1}$ of $\pintss$ that has no immediate predecessor in $<$ but also such that $f\ppt{1} \neq (1,1)$ so that $f\ppt{1}$ does have an immediate predecessor in $\prec$.
    So let $(u_0, v_0)$ be the immediate predecessor of $f\ppt{1}$ in $\prec$ and set $\ppt{0} = \inv{f}(u_0,v_0)$ so that $(u_0,v_0) = f\ppt{0}$.
    Then of course $\ppt{0} < \ppt{1}$ since $f\ppt{0} = (u_0,v_0) \prec f\ppt{1}$ and $f$ preserves order.
    But since $\ppt{1}$ has no immediate predecessor in $<$, there is a point $\ppt{2}$ such that $\ppt{0} < \ppt{2} < \ppt{1}$.
    We then have that $(u_0,v_0) = f\ppt{0} \prec f\ppt{2} \prec f\ppt{1}$ since $f$ preserves order, which is a contradiction since $(u_0,v_0)$ is the immediate predecessor of $f\ppt{1}$.
    So it must be that no such order-preserving $f$ exists and hence the two orders do not have the same order type.
  }
}

\exercise{13}{
  Prove the following:
  \begin{thrm*}
    If an ordered set $A$ has the least upper bound property, then it has the greatest lower bound property.
  \end{thrm*}
}
\sol{
  \qproof{
    Suppose that $A_0$ is any nonempty subset of $A$ that is bounded below so that $b$ is a lower bound of $A_0$.
    Let $B_0$ be the set of lower bounds of $A_0$ so that $B_0$ is nonempty since $b \in B_0$.
    Since $A_0$ is nonempty there is an $a \in A_0$.
    Now, for any $x \in B_0$ we have that $x$ is a lower bound of $A_0$ so that $x \leq a$, which shows that $a$ is an upper bound of $B_0$.
    Hence $B_0$ is a nonempty subset of $A$ that is bounded above, and so has a least upper bound $c$ since $A$ has the least upper bound property.
    We claim that $c$ is also the greatest lower bound of $A_0$.

    Consider any $x \in A_0$ and any $y \in B_0$.
    Then $y$ is a lower bound of $A_0$ so that $y \leq x$ since $x \in A_0$.
    Since $y \in B_0$ was arbitrary, this shows that $x$ is an upper bound of $B_0$.
    Thus we have $c \leq x$ since $c$ is the least upper bound of $B_0$.
    Since $x \in A_0$ was arbitrary, this shows that $c$ is a lower bound of $A_0$.
    If $y$ is any other lower bound then $y \in B_0$ so that $y \leq c$ since $c$ is an upper bound of $B_0$.
    Since $y$ was an arbitrary lower bound, this shows that in fact $c$ is the greatest lower bound of $A_0$.
    Hence $A$ also has the greatest lower bound property since the nonempty subset $A_0$ was arbitrary.
  }
}

\exercise{14}{
  If $C$ is a relation on a set $A$, define a new relation $D$ on $A$ by letting $(b,a) \in D$ if $(a,b) \in C$.
  \eparts{
  \item Show that $C$ is symmetric if and only if $C = D$.
  \item Show that if $C$ is an order relation, $D$ is also an order relation.
  \item Prove the converse of the theorem in Exercise~13.
  }
}
\sol{
  (a)
  \qproof{
    $(\imp)$ Suppose that $C$ is symmetric.
    Then we simply have
    \ali{
      (a,b) \in C &\bic (b,a) \in C & \text{(since $C$ is symmetric)} \\
      &\bic (a,b) \in D \,, & \text{(by definition)}
    }
    which shows that $C = D$.

    $(\pmi)$ Now suppose that $C = D$ and consider any $(a,b) \in C$.
    Then $(b,a) \in D$ by definition.
    Hence also $(b,a) \in C$ since $C = D$, which shows that $C$ is symmetric.
  }

  (b)
  \qproof{
    Suppose that $C$ is an order relation.
    Since clearly $D$ is a relation on $A$, we need only show that it has the three required properties:

    (Comparability) Consider any distinct $a,b \in A$ so that $aCb$ or $bCa$ since $C$ has comparability.
    Hence either $bDa$ or $aDb$, respectively, by definition so that $a$ and $b$ are comparable in $D$ as well.

    (Nonreflexivity) Consider any $a \in A$.
    Then $(a,a) \notin C$ since it is nonreflexive, thus also $(a,a) \notin D$ since, if it were, it would also be that $(a,a) \in C$ by definition.
    Hence $D$ is also nonreflexive since $a$ was arbitrary.

    (Transitivity) Suppose that $aDb$ and $bDc$.
    Then by definition we have $bCa$ and $cCb$.
    That is, $cCb$ and $bCa$ so that $cCa$ since $C$ is transitive.
    Therefore $aDc$ by definition, which shows that $D$ is also transitive.
  }

  (c)
  The converse follows from an argument directly analogous to the proof of Exercise~3.13, which we give here for completeness.
  \qproof{
    Suppose that $A$ has the greatest lower bound property and that $A_0$ is any nonempty subset of $A$ that is bounded above so that $b$ is an upper lower bound of $A_0$.
    Let $B_0$ be the set of upper bounds of $A_0$ so that $B_0$ is nonempty since $b \in B_0$.
    Since $A_0$ is nonempty there is an $a \in A_0$.
    Now, for any $x \in B_0$ we have that $x$ is an upper bound of $A_0$ so that $a \leq x$, which shows that $a$ is a lower bound of $B_0$.
    Hence $B_0$ is a nonempty subset of $A$ that is bounded below, and so has a greatest lower bound $c$ since $A$ has the greatest lower bound property.
    We claim that $c$ is also the least upper bound of $A_0$.

    Consider any $x \in A_0$ and any $y \in B_0$.
    Then $y$ is an upper bound of $A_0$ so that $x \leq y$ since $x \in A_0$.
    Since $y \in B_0$ was arbitrary, this shows that $x$ is a lower bound of $B_0$.
    Thus we have $x \leq c$ since $c$ is the greatest lower bound of $B_0$.
    Since $x \in A_0$ was arbitrary, this shows that $c$ is an upper bound of $A_0$.
    If $y$ is any other upper bound then $y \in B_0$ so that $c \leq y$ since $c$ is a lower bound of $B_0$.
    Since $y$ was an arbitrary upper bound, this shows that in fact $c$ is the lest upper bound of $A_0$.
    Hence $A$ also has the lest upper bound property since the nonempty subset $A_0$ was arbitrary.
  }
}

\renewcommand\ppt[1]{x_{#1} \times y_{#1}}
\exercise{15}{
  Assume that the real line has the least upper bound property.
  \eparts{
  \item Show that the sets
    \ali{
      [0,1] &= \braces{x \where 0 \leq x \leq 1} \,, \\
      \clop{0,1} &= \braces{x \where 0 \leq x < 1}
    }
    have the least upper bound property.
  \item Does $[0,1] \times [0,1]$ in the dictionary order have the least upper bound property?
    What about $[0,1] \times \clop{0,1}$?
    What about $\clop{0,1} \times [0,1]$?
  }
}
\sol{
  (a)
  We show this for both sets simultaneously as their proofs are nearly identical.
  We note the minor differences in parentheses.
  \qproof{
    Let $A = [0,1]$ (or $A = \clop{0,1}$) and consider any nonempty subset $A_0$ of $A$ that is bounded above in $A$.
    So suppose that $b$ is an upper bound of $A_0$ in $A$ so that $0 \leq b \leq 1$ (or $0 \leq b < 1$).
    Now, of course $A_0$ has a least upper bound $c$ in $\reals$ since it is also a nonempty subset of $\reals$ that is bounded above.
    Obviously $c \leq b$ since it is the \emph{least} upper bound.
    For any $x \in A_0$ we of course have that $x \in A$ so that $0 \leq x \leq 1$ (or $0 \leq x < 1$), and clearly $x \leq c$ since it is an upper bound of $A_0$.
    Thus we have $0 \leq x \leq c \leq b \leq 1$ (or $0 \leq x \leq c \leq b <1$) so that $c \in A$.
    Hence $A_0$ has a least upper bound in $A$ as desired.
  }

  (b)
  First we claim that both $[0,1] \times [0,1]$ and $\clop{0,1} \times [0,1]$ have the least upper bound property.
  We show this for both sets simultaneously as their proofs are identical.
  \qproof{
    Let $X = [0,1] \times [0,1]$ (or $X = \clop{0,1} \times [0,1]$).
    Suppose that $A$ is a nonempty subset of $X$ that is bounded above, and that $\ppt{1}$ is an upper bound of $A$ in $X$.
    Then of course we have $x_1 \in [0,1]$ (or $x_1 \in \clop{0,1}$).
    Define $A_x = \braces{x \where \ppt{} \in A}$ so that clearly $A_x \ss [0,1]$ (or $A_x \ss \clop{0,1}$).
    It then follows that $x \leq x_1$ for any $x \in A_x$ since $\ppt{1}$ is an upper bound of $A$ in the dictionary order.
    Also clearly $A_x$ is nonempty since $A$ is.
    Thus $A_x$ is a nonempty subset of $[0,1]$ (or $\clop{0,1}$) that is bounded above (by $x_1$) so that it has a least upper bound $a$ by what was shown in part~(a).

    Now, if $a \notin A_x$ then set $b = 0$.
    Otherwise define $A_y = \braces{y \where a \times y \in A}$.
    Then, since $a \in A_x$, there is a $y \in [0,1]$ where $a \times y \in A$, which shows that $y \in A_y$ and hence $A_y \neq \es$.
    We also clearly have that $A_y \ss [0,1]$ so that $A_y$ is bounded above by $1$.
    Then $A_y$ has a least upper bound $b$, again by what was shown in part~(a).
    In either case we assert that $a \times b$ is the least upper bound of $A$ in $X$ in the dictionary order.

    First, it is obvious that $a \times b \in X$ based on how $a$ and $b$ were defined.
    Consider any $\ppt{} \in A$ so that $x \in A_x$, and hence $x \leq a$ since $a$ is the least upper bound of $A_x$.
    If $x < a$ then of course $\ppt{} \leq a \times b$, so assume that $x = a$.
    Then $a = x \in A_x$ so that $b$ was defined as the least upper bound of $A_y$.
    Then we have that $y \in A_y$ since $\ppt{} = a \times y \in A$, and thus $y \leq b$ since $b$ is the least upper bound of $A_y$.
    This shows that $\ppt{} \leq a \times b$ so that we  have shown that $a \times b$ is an upper bound of $A$ since $\ppt{}$ was arbitrary.

    To show that it is the \emph{least} upper bound consider any $\ppt{0} \in X$ where $\ppt{0} < a \times b$.
    If $x_0 < a$ then there must be an $x \in A_x$ where $x_0 < x$ since $x_0$ cannot be an upper bound of $A_x$ since $x_0 < a$ and $a$ is the least upper bound of $A_x$.
    Then, since $x \in A_x$, there is a $y \in [0,1]$ where $\ppt{} \in A$.
    Hence $\ppt{0} < \ppt{}$ since $x_0 < x$ so that $\ppt{0}$ is not an upper bound of $A$.
    On the other hand if $x_0 = a$ we must have $y_0 < b$ so that it cannot be that $b = 0$, and hence it must be that $a \in A_x$.
    Then there must be a $y \in A_y$ where $y_0 < y$ since $y_0$ cannot be an upper bound of $A_y$ since $y_0 < b$ and $b$ is the least upper bound of $A_y$.
    Hence $a \times y \in A$, $x_0 = a$, and $y_0 < y$ so that $\ppt{0} < a \times y \in A$, which shows that $\ppt{0}$ is not an upper bound of $A$.
    Thus in either case $\ppt{0}$ is not an upper bound of $A$, which shows that $a \times b$ is the least upper bound since $\ppt{0} < a \times b$ was arbitrary.

    This of course completes the proof that $A$ has a least upper bound, which shows that $X$ has the least upper bound property since $A$ was an arbitrary nonempty subset.
  }

  We also claim that $[0,1] \times \clop{0,1}$ does \emph{not} have the least upper bound property.
  \qproof{
    Let $X = [0,1] \times \clop{0,1}$.
    Consider the set $A = \braces{0} \times \clop{0,1}$, which is clearly a nonempty subset of $X$.
    This subset also obviously has an upper bound in $X$ in the dictionary order, for example the point $1 \times 0$.\
    So let $\ppt{1}$ be any upper bound of $A$ in $X$ and suppose for the moment that $x_1 = 0$.
    Then $y_1 \in \clop{0,1}$ so that $0 \leq y_1 < 1$, but then there is a $y_0 \in \reals$ such that $0 \leq y_1 < y_0 < 1$.
    Then $y_0 \in \clop{0,1}$ so that $x_1 \times y_0 = 0 \times y_0 \in A$, but also $\ppt{1} < x_1 \times y_0$ so that $\ppt{1}$ cannot be an upper bound of $A$.
    So it must be that in fact $x_1 \neq 0$ and hence $x_1 > 0$.
    Now let $x_0 = x_1/2 > 0$ and $y_0 = 0$ so that clearly $\ppt{} < \ppt{0} < \ppt{1}$ for any $\ppt{} \in A$ since $x = 0 < x_1/2 = x_0 < x_1$.
    This shows that $\ppt{0}$ is still an upper bound of $A$ but that $\ppt{0} < \ppt{1}$.
    Since $\ppt{1}$ was an arbitrary upper bound of $A$, this proves that $A$ can have no \emph{least} upper bound!
  }
}
